---
title: "R Bridge Week 3 Assignment"
author: "Rob Hodde"
date: 11/15/2015
output: html_document
---


***

### Problem Statement:

Very often, we're tasked with taking data in one form and transforming it for easier downstream analysis.  We will
spend several weeks in the Fall 607 class on tidying and transformation operations. Some of this work could be
done in SQL or R (or Python or.). Here, you are asked to use R-you may use base functions or packages as you like.  


You need to:  

 - create a data frame with a subset of the columns (and if you like rows) in the dataset.
 
 - include the column that indicates edible or poisonous and three or four other columns
 
 - add meaningful column names and replace the abbreviations used in the data 
  
Your deliverable is the R code to perform these transformation tasks.

You should also have the original data file accessible through your code - for example, stored in a github repository and referenced in your code.  



***
  
### Solution:

First we will make a factorial function to be used in the "n chooses r" function:

```{r}

fn_calc_factorial <- function(n) {
  
  iproduct <- 1 # set initial value
  
  # build up factorial value in a loop
  for(i in 1:n) {
   iproduct <- iproduct * i  # multiply the product by the new value to build up the factorial
  }
  
  return (iproduct)
}

```

  


Next we will create the "n choose r" function:

```{r}

fn_calc_choose <- function(n,r) {
  
  # we are performing this action:  n! / (( n - r)! * r!)
  ichoose <- fn_calc_factorial(n) / (fn_calc_factorial(n - r) * fn_calc_factorial(r))
  
  return (ichoose)  
}
   
```

***
  
### Test Cases

**Test 1:**  If three flavors of ice cream are available, how many different double scoop cones can be made?

```{r}

test_1 <- fn_calc_choose(30, 2)
test_1

```

*Answer:* `r fn_calc_choose(3, 2)`


**Test 2:**  How many different assemblies can we build using four components if we have 15 componenents available?

```{r}

test_2 <- fn_calc_choose(15, 4)
test_2

```

*Answer:* `r fn_calc_choose(15, 4)`


**Test 3:**  Unit Test of factorial function:

```{r}

 ifactorial <- fn_calc_factorial(4)  # answer should be 1 x 2 x 3 x 4 = 24.
 ifactorial
```

*Answer:* `r fn_calc_factorial(4)`




