---
title: "R Bridge Week 3 Assignment"
author: "Rob Hodde"
date: 11/15/2015
output: html_document
---


***

### Problem Statement:

Very often, we're tasked with taking data in one form and transforming it for easier downstream analysis.  We will
spend several weeks in the Fall 607 class on tidying and transformation operations. Some of this work could be
done in SQL or R (or Python or.). Here, you are asked to use R-you may use base functions or packages as you like.  


You need to:  

 - create a data frame with a subset of the columns (and if you like rows) in the dataset.
 
 - include the column that indicates edible or poisonous and three or four other columns
 
 - add meaningful column names and replace the abbreviations used in the data 
  
Your deliverable is the R code to perform these transformation tasks.

You should also have the original data file accessible through your code - for example, stored in a github repository and referenced in your code.  



***
  
### Solution:

First we will import a few rows of the ICS mushroom data set, excerpted from https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data.

```{r}

sURL <- "https://raw.githubusercontent.com/LovinSpoonful/Week3-R/62d6aed42bdf8869a8c692c0c0108782703de249/MushData.csv"
dfMushRoom <- read.table(file = sURL, header = FALSE, sep = ",", stringsAsFactors = FALSE)

head(dfMushRoom)  # verification


```
  


The first column in the dataset is the calculated classification (edible / poisonous); the other 22 are the observed mushroom characteristics.


The data is very hard to understand because the column headers are just cryptic codes.  Let's replace them with real descriptions from the source study, which we find here:
https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names.


```{r}

# todo: find a way not to hard code these literals
cAttributeNames <- c("edible","cap-shape", "cap-surface", "cap-color", "bruises", "odor", "gill-attachment", 
          "gill-spacing", "gill-size", "gill-color", "stalk-shape", "stalk-root", 
          "stalk-surface-above-ring", "stalk-surface-below-ring", "stalk-color-above-ring", 
          "stalk-color-below-ring", "veil-type", "veil-color", "ring-number", "ring-type", 
          "spore-print-color", "population", "habitat")

colnames(dfMushRoom) <- cAttributeNames

head(dfMushRoom)

```
  

  

Next we will focus on the most important attributes for determining edibility, according to the commentary below from the source study - https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names -

  
  	P_1) odor=NOT(almond.OR.anise.OR.none)
  	     120 poisonous cases missed, 98.52% accuracy
  
  	P_2) spore-print-color=green
  	     48 cases missed, 99.41% accuracy
           
  	P_3) odor=none.AND.stalk-surface-below-ring=scaly.AND.
  	          (stalk-color-above-ring=NOT.brown) 
  	     8 cases missed, 99.90% accuracy
           
  	P_4) habitat=leaves.AND.cap-color=white
  	         100% accuracy
	         
    
     odor:                     almond=a,anise=l,creosote=c,fishy=y,foul=f,
                               musty=m,none=n,pungent=p,spicy=s
     
     spore-print-color:        black=k,brown=n,buff=b,chocolate=h,green=r,
                               orange=o,purple=u,white=w,yellow=y
     
     stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s
     
     stalk-color-above-ring:   brown=n,buff=b,cinnamon=c,gray=g,orange=o,
                               pink=p,red=e,white=w,yellow=y
     
     habitat:                  grasses=g,leaves=l,meadows=m,paths=p,
                               urban=u,waste=w,woods=d     
    
     cap-color:                brown=n,buff=b,cinnamon=c,gray=g,green=r,
                               pink=p,purple=u,red=e,white=w,yellow=y


Here are the main characteristics that determine if you will die after you eat the mushroom:

```{r}

# todo: don't hard code these literals -  get them from the Class Codes headers
dfClassifier <- subset(dfMushRoom, select = c("edible","odor","spore-print-color","stalk-surface-below-ring","stalk-color-above-ring","habitat","cap-color"))
dfClassifier

```


That looks a little better but we still have a bunch of cryptic data in the values.  Let's use a lookup table to cross references the mushroom attribute values to human-friendly terms:

```{r}

#  get the key classifiers, value codes and value decriptions
sURL =   "https://raw.githubusercontent.com/LovinSpoonful/Week3-R/1646a130a20704f936a5da89cd03bcddef14a694/AttributeValues.csv"
dfClassCodes <- read.table(file = sURL, header = TRUE, sep = ",", stringsAsFactors = FALSE)

colnames(dfClassCodes)
nrow(dfClassCodes)
dfClassCodes

```


We use a nested loop to overwrite the cryptic values with human friendly terms:

```{r}

for (i in 1:nrow(dfClassCodes))
{
  sAttribute <- dfClassCodes[i,1]
  sCode <- dfClassCodes[i,2]
  sDescription <- dfClassCodes[i,3]
  
  for (j in 1:nrow(dfClassifier))
  {
    if(dfClassifier[j,sAttribute] == sCode)
    {
      dfClassifier[j,sAttribute] <- sDescription
    }
  }
}

dfClassifier

```



