---
title: "R Bridge Week 3 Assignment"
author: "Rob Hodde"
date: 11/15/2015
output: html_document
---


***

### Problem Statement:

Very often, we're tasked with taking data in one form and transforming it for easier downstream analysis.  We will
spend several weeks in the Fall 607 class on tidying and transformation operations. Some of this work could be
done in SQL or R (or Python or.). Here, you are asked to use R-you may use base functions or packages as you like.  


You need to:  

 - create a data frame with a subset of the columns (and if you like rows) in the dataset.
 
 - include the column that indicates edible or poisonous and three or four other columns
 
 - add meaningful column names and replace the abbreviations used in the data 
  
Your deliverable is the R code to perform these transformation tasks.

You should also have the original data file accessible through your code - for example, stored in a github repository and referenced in your code.  



***
  
### Solution:

First we will import the mushroom data:

```{r}

sURL <- "https://raw.githubusercontent.com/LovinSpoonful/Week3-R/62d6aed42bdf8869a8c692c0c0108782703de249/MushData.csv"
dfMushRoom <- read.table(file = sURL, header = FALSE, sep = ",", stringsAsFactors = FALSE)

head(dfMushRoom)  # verification


```
  

  

Twenty two mushroom attributes are listed in https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names.

The first column in the dataset is the classification (edible / poisonous), the rest are the mushroom characteristics.


Let's assign the dataset column headers to the dataframe column headers:


```{r}

# todo: find a way not to hard code these literals
cAttributeNames <- c("edible","cap-shape", "cap-surface", "cap-color", "bruises", "odor", "gill-attachment", 
          "gill-spacing", "gill-size", "gill-color", "stalk-shape", "stalk-root", 
          "stalk-surface-above-ring", "stalk-surface-below-ring", "stalk-color-above-ring", 
          "stalk-color-below-ring", "veil-type", "veil-color", "ring-number", "ring-type", 
          "spore-print-color", "population", "habitat")

colnames(dfMushRoom) <- cAttributeNames

head(dfMushRoom)

```
  

  

Next we will focus on the most important attributes for determining edibility. 

Again, from https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names -

  
  	P_1) odor=NOT(almond.OR.anise.OR.none)
  	     120 poisonous cases missed, 98.52% accuracy
  
  	P_2) spore-print-color=green
  	     48 cases missed, 99.41% accuracy
           
  	P_3) odor=none.AND.stalk-surface-below-ring=scaly.AND.
  	          (stalk-color-above-ring=NOT.brown) 
  	     8 cases missed, 99.90% accuracy
           
  	P_4) habitat=leaves.AND.cap-color=white
  	         100% accuracy
	         
    
     odor:                     almond=a,anise=l,creosote=c,fishy=y,foul=f,
                               musty=m,none=n,pungent=p,spicy=s
     
     spore-print-color:        black=k,brown=n,buff=b,chocolate=h,green=r,
                               orange=o,purple=u,white=w,yellow=y
     
     stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s
     
     stalk-color-above-ring:   brown=n,buff=b,cinnamon=c,gray=g,orange=o,
                               pink=p,red=e,white=w,yellow=y
     
     habitat:                  grasses=g,leaves=l,meadows=m,paths=p,
                               urban=u,waste=w,woods=d     
    
     cap-color:                brown=n,buff=b,cinnamon=c,gray=g,green=r,
                               pink=p,purple=u,red=e,white=w,yellow=y


Let's make a subset of classifier attribute names:

```{r}

# todo: find a way not to hard code these literals
dfClassifier <- subset(dfMushRoom, select = c("edible","odor","spore-print-color","stalk-surface-below-ring","stalk-color-above-ring","habitat","cap-color"))
colnames(dfClassifier)

dfClassifier




#  get the key classifiers, value codes and value decriptions

sURL = "https://raw.githubusercontent.com/LovinSpoonful/Week3-R/74daca35112f48ccc47d6ee33d8c9c3458c3cfb4/AttributeValues.csv"
dfClassCodes <- read.table(file = sURL, header = TRUE, sep = ",", stringsAsFactors = FALSE)

colnames(dfClassCodes)
nrow(dfClassCodes)
dfClassCodes


for (i in 1:nrow(dfClassCodes))
{
  sAttribute <- dfClassCodes[i,1]
  sCode <- dfClassCodes[i,2]
  sDescription <- dfClassCodes[i,3]
  
  # i'm trying to say "check every row in dfClassifier, column sAttribute - if the code is sCode, change it to sDescription""
  # dfClassifier[1:nrow(dfClassifier),sAttribute]
  
  for (j in 1:nrow(dfClassifier))
  {
    print(dfClassifier[j,sAttribute])
    if(dfClassifier[j,sAttribute] == sCode)
    {
      dfClassifier[j,sAttribute] <- sDescription
    }
    print(dfClassifier[j,sAttribute])
  }
}

dfClassifier



```



