---
title: "R Bridge Week 3 Assignment"
author: "Rob Hodde"
date: 11/15/2015
output: html_document
---


***

### Problem Statement:

Very often, we're tasked with taking data in one form and transforming it for easier downstream analysis.  We will
spend several weeks in the Fall 607 class on tidying and transformation operations. Some of this work could be
done in SQL or R (or Python or.). Here, you are asked to use R-you may use base functions or packages as you like.  


You need to:  

 - create a data frame with a subset of the columns (and if you like rows) in the dataset.
 
 - include the column that indicates edible or poisonous and three or four other columns
 
 - add meaningful column names and replace the abbreviations used in the data 
  
Your deliverable is the R code to perform these transformation tasks.

You should also have the original data file accessible through your code - for example, stored in a github repository and referenced in your code.  



***
  
### Solution:

First we will import the mushroom data:

```{r}

sURL <- "https://raw.githubusercontent.com/LovinSpoonful/Week3-R/62d6aed42bdf8869a8c692c0c0108782703de249/MushData.csv"
dfMushRoom <- read.table(file = sURL, header = FALSE, sep = ",", stringsAsFactors = FALSE)

ncol(dfMushRoom)
nrow(dfMushRoom)

```
  

  

Twenty two mushroom attributes are listed in https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names.

  
Since there are twenty three columns in the data set, I suspect the 23rd column indicates if the mushroom is edible.


Let's assign the mushroom attributes as the dataframe column header names:


```{r}

cDim <- c("cap-shape", "cap-surface", "cap-color", "bruises", "odor", "gill-attachment", 
          "gill-spacing", "gill-size", "gill-color", "stalk-shape", "stalk-root", 
          "stalk-surface-above-ring", "stalk-surface-below-ring", "stalk-color-above-ring", 
          "stalk-color-below-ring", "veil-type", "veil-color", "ring-number", "ring-type", 
          "spore-print-color", "population", "habitat", "edible")

colnames(dfMushRoom) <- cDim

```
  

  

Next we will focus on the most important attributes for determining edibility. 

Again, from https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names -

  
  	P_1) odor=NOT(almond.OR.anise.OR.none)
  	     120 poisonous cases missed, 98.52% accuracy
  
  	P_2) spore-print-color=green
  	     48 cases missed, 99.41% accuracy
           
  	P_3) odor=none.AND.stalk-surface-below-ring=scaly.AND.
  	          (stalk-color-above-ring=NOT.brown) 
  	     8 cases missed, 99.90% accuracy
           
  	P_4) habitat=leaves.AND.cap-color=white
  	         100% accuracy
	         
    
    
     odor:                     almond=a,anise=l,creosote=c,fishy=y,foul=f,
                               musty=m,none=n,pungent=p,spicy=s
     
     spore-print-color:        black=k,brown=n,buff=b,chocolate=h,green=r,
                               orange=o,purple=u,white=w,yellow=y
     
     stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s
     
     stalk-color-above-ring:   brown=n,buff=b,cinnamon=c,gray=g,orange=o,
                               pink=p,red=e,white=w,yellow=y
     
     habitat:                  grasses=g,leaves=l,meadows=m,paths=p,
                               urban=u,waste=w,woods=d                                 
                                  
                                  
 


```{r}

dfEdible <- subset(dfMushRoom, select = c("odor","spore-print-color","stalk-surface-below-ring","stalk-color-above-ring","habitat","cap-color","edible"))

dfEdible







```



